Fourmile state variables

### 2025-05-06 state_vars.Rmd
looking at physical properties that are impacted by burn params

both_hi <- bind_rows(statn_hi, grid_hi) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  filter(year == 2016,
         month %in% c(4,5,6))

@@

>>>>

both_hi <- bind_rows(statn_hi, grid_hi) %>%
  mutate(year = year(date),
         month = month(date)) %>%
  filter(year == 2016) #
         #month %in% c(4,5,6))

@@

both <- bind_rows(unburn, burn)

# interception
inter <- both %>%
  dplyr::select(date, type, intercStorage)

int_plot <- ggplot(inter,
       aes(x=date,,
           y=intercStorage,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

ggplotly(int_plot)

# swe
swe <- both %>%
  dplyr::select(date, type, snowTotSWE)

ggplot(swe,
       aes(x=date,,
           y=snowTotSWE,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# swe
p <- both %>%
  dplyr::select(date, type, precip) %>%
  group_by(type) %>%
  mutate(p_cumu = cumsum(precip))

ggplot(p,
       aes(x=date,
           y=p_cumu,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# soil sat
soilsat <- both %>%
  dplyr::select(date, type, soilSat)

ggplot(soilsat,
       aes(x=date,
           y=soilSat,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# surface runoff
runoff <- both %>%
  dplyr::select(date, type, outRD1)

ggplot(runoff,
       aes(x=date,,
           y=outRD1,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# interflow
interflow <- both %>%
  dplyr::select(date, type, outRD2) %>%
  mutate(interflow_cum = cumsum(outRD2))

ggplot(interflow,
       aes(x=date,,
           y=interflow_cum,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

gw <- both %>%
  dplyr::select(date, type, outRG1) %>%
  mutate(gw_cumu = cumsum(outRG1))

ggplot(gw,
       aes(x=date,,
           y=gw_cumu,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# dps
dps <- both %>%
  dplyr::select(date, type, actDPS)

ggplot(dps,
       aes(x=date,,
           y=actDPS,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# mps
mps <- both %>%
  dplyr::select(date, type, soilSatMPS)

ggplot(mps,
       aes(x=date,,
           y=soilSatMPS,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# lps
lps <- both %>%
  dplyr::select(date, type, soilSatLPS)

ggplot(lps,
       aes(x=date,,
           y=soilSatLPS,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

# soil sat
soilsat <- both %>%
  dplyr::select(date, type, soilSat)

ggplot(soilsat,
       aes(x=date,,
           y=soilSat,
           color=type)) +
  geom_line() +
  theme_bw(base_size=20)

## make the actual plot
p_plot <- ggplot(p, aes(x = date,
                        y = p_cumu, 
                        color = type)) +
  geom_line(size=1, alpha=0.8) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Cumulative precipitation (mm)') +
  theme_bw(base_size = 18) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

p_plot

inter <- inter %>%
  mutate(type = factor(type, levels = c("unburn", "burn")))

int_plot <- ggplot(inter, aes(x = date,
                        y = intercStorage, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Interception (mm)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

int_plot

soilsat_plot <- ggplot(soilsat, aes(x = date,
                        y = soilSat, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Soil saturation') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

soilsat_plot

interflow_plot <- ggplot(interflow, aes(x = date,
                        y = interflow_cum, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Interflow (L)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

interflow_plot

gw_plot <- ggplot(gw, aes(x = date,
                        y = gw_cumu, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'Shallow groundwater (L)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

gw_plot

swe_plot <- ggplot(swe, aes(x = date,
                        y = snowTotSWE, 
                        color = type)) +
  geom_line(size=1, alpha=0.7) +
    scale_color_manual(values = c("burn" = "red", 
                                  "unburn" = "black")) +
  scale_x_date(date_labels = "%b") +
  labs(x = element_blank(),
       y = 'SWE (mm)') +
  theme_bw(base_size = 20) +
  theme(legend.position = 'bottom',
    #legend.position = c(0.24, 0.852), 
    legend.text = element_text(size = 10),
    legend.title = element_blank(),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    legend.key.width = unit(1, "cm")
  )

swe_plot

legend <- cowplot::get_plot_component(p_plot, "guide-box", return_all = TRUE)[[3]]

p_plot <- p_plot + theme(legend.position = "none")
int_plot <- int_plot + theme(legend.position = "none")
soilsat_plot <- soilsat_plot + theme(legend.position = "none")
interflow_plot <- interflow_plot + theme(legend.position = "none")
gw_plot <- gw_plot + theme(legend.position = "none")
swe_plot <- swe_plot + theme(legend.position = "none")

plots <- plot_grid(p_plot,
          int_plot,
          swe_plot,
          soilsat_plot,
          interflow_plot,
          gw_plot,
          ncol = 3)

final_plot <- plot_grid(
  plots,
  legend,
  ncol = 1,
  rel_heights = c(1, 0.07)  # Adjust height of legend
)

final_plot

ggsave('./figures/burn_unburn_eg.png')

soilsat_mean <- soilsat %>% 
  group_by(type) %>% 
  summarize(mean_ss = mean(soilSat))

soilsat_mean

intercep_mean <- inter %>% 
  group_by(type) %>% 
  summarize(mean_inter = mean(intercStorage))

intercep_mean
